Concurrent Programming in Harmony

### 2021-12-26 book/paper.tex
Fixed some more formatting
Latest

@@

\glsadd{state machine replication}%
\index{state machine replication}%
\index{determinism}%
Besides availability, it is usually important that the replicated
service acts as if it were a single one.  This requires that the
replicas of the service coordinate their actions.
The \emph{Replicated State Machine Approach}~\cite{Lam78,S90} is a
general approach to do just this.
First, you model your service as a deterministic state
machine.  The replicas each run a copy of the state machine, started
in the same state.  As long as the replicas handle the same inputs
in the same order, determinism guarantees that they produce the same
outputs in the same order.
The trick then is to ensure that all replicas handle the
same requests in the same order and to do so in a way that continues
to work even if some strict subset of replicas crash.

\emph{Chain Replication}~\cite{vRS04} is such a replication protocol.
In Chain Replication, the replicas are organized in a linear chain.
Each replica monitors its direct predecessor in the chain.  If a replica has
no predecessors, we call it the \emph{head}.  The head may change over
time, as a head may crash and replaced by another, but at any point in
time there is only one head
(and only one \emph{tail}, possibly the same replica).
We model the state machine as a \emph{history}: a sequence of operations.
Only the head is allowed to introduce new updates.
When it does so, it advertises the new update along with its position
in the history.  When the direct successor receives the next update
it is expecting, it appends the update to its history and likewise
advertises the new update.  When the update is applied to the history
of the tail replica (the replica that has no live successors),
then the update is considered final.

@@

\autoref{fig:chain} shows an implemenation of chain replication.
The code starts \texttt{NREPLICAS} \texttt{replica} threads numbered
0 through $\mathtt{NREPLICAS} - 1$, with initially 0 being the head
and $\mathtt{NREPLICAS} - 1$ being the tail.  At least one of the replicas 
must be immortal---the code selects one of them although none of the
replicas is required to fail during execution.
Of course, a replica does not know whether it is immortal
or not in practice---it should just assume so.
The immortality of one of the replicas is only used for modeling the system.

Each replica maintains its history and a set of its predecessors.
It then loops until either it fails or it has applied all
<{NOPS}> operations to its history.
$\mathbf{choose}(\{ \mathit{immortal}, <{True}> \})$ can only
evaluate to <{False}> in case $<{immortal}>$ is false; otherwise
it will always evaluate to <{True}>.
What it does within the loop depends on whether it is the head
(has no alive predecessors) or not.
Because failure detection is accurate, at most one replica can think
it is the head at any time (and, if so, it is in fact the head).
Moreover, when it has
detected all its predecessors having failed, eventually some replica
thinks it is the head.

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
<{tas(lk)}> & test-and-set on <{!lk}> \\
\hline
<{cas}(ptr, old, new)}> & compare-and-swap on <{!ptr}> \\
\hline
<{BinSem(v)}> & return a binary semaphore initialized to <{v}> \\
\hline
<{Lock()}> & return a binary semaphore initialized to <{False}> \\
\hline
<{acquire(bs)}> & acquire binary semaphore <{!bs}> \\
\hline
<{release(bs)}> & release binary semaphore <{!bs}> \\
\hline
<{Condition()}> & return a condition variable \\
\hline
<{wait(c, lk)}> & wait on condition variable <{!c}>
and lock <{lk}> \\ \hline
<{notify(c)}> & notify a thread waiting on condition variable <{!c}> \\
\hline
<{notifyAll(c)}> & notify all threads waiting on condition variable <{!c}> \\
\hline
<{Semaphore(cnt)}> & return a counting semaphore initialized to <{cnt}> \\
\hline
<{P(sema)}> & procure <{!sema}>  \\
\hline
<{V(sema)}> & vacate <{!sema}>  \\
\hline
<{Queue}>() & return a synchronized queue object \\
\hline
<{get(q)}> & return next element of <{q}>, blocking if empty \\
\hline
<{put(q, item)}> & add <{item}> to <{a}> \\
\hline
\end{tabular}

>>>>

<{:const:tas}>
<{:const:cas}>
<{:const:BinSem}>
<{:const:Semaphore}>

@@

\glsadd{state machine replication}%
\index{state machine replication}%
\index{determinism}%
Besides availability, it is usually important that the replicated
service acts as if it were a single one.  This requires that the
replicas of the service coordinate their actions.
The \emph{Replicated State Machine Approach}~\cite{Lam78,S90} is a
general approach to do just this.
First, you model your service as a deterministic state
machine.  The replicas each run a copy of the state machine, started
in the same state.  As long as the replicas handle the same inputs
in the same order, determinism guarantees that they produce the same
outputs in the same order.
Therefore, the trick is to ensure that all replicas handle the
same requests in the same order and to do so in a way that continues
to work even if some strict subset of replicas crash.

\emph{Chain Replication}~\cite{vRS04} is such a replication protocol.
In Chain Replication, the replicas are organized in a linear chain.
Each replica monitors its direct predecessor in the chain.  If a replica has
no predecessors, we call it the \emph{head}.  The head may change over
time, as a head may crash and replaced by another, but at any point in
time there is only one head
(and only one \emph{tail}, possibly the same replica).

We model the state machine as a \emph{history}: a sequence of operations.
Only the head is allowed to introduce new updates.
When it does so, it advertises the new update along with its position
in the history.  When the direct successor receives the next update
it is expecting, it appends the update to its history and likewise
advertises the new update.  When the update is applied to the history
of the tail replica (the replica that has no live successors),
then the update is considered final.

@@

\autoref{fig:chain} shows an implemenation of chain replication.
The code starts \texttt{NREPLICAS} \texttt{replica} threads numbered
0 through $\mathtt{NREPLICAS} - 1$, with initially 0 being the head
and $\mathtt{NREPLICAS} - 1$ being the tail.  At least one of the replicas 
must be immortal---the code selects one of them although none of the
replicas is required to fail during execution.
Of course, a replica does not know whether it is immortal
or not in practice---it should just assume that it is.
The immortality of one of the replicas is only used for modeling
the assumptions we make about the system.

Each replica maintains its history and a set of its predecessors.
It then loops until either it fails or it has applied all
<{NOPS}> operations to its history.
<{choose({ immortal, True })}> can only
evaluate to <{False}> in case <{immortal = False}>; otherwise
it will always evaluate to <{True}>.
What it does within the loop depends on whether it is the head
(has no alive predecessors) or not.
Because failure detection is accurate, at most one replica can think
it is the head at any time (and, if so, it is in fact the head).
Moreover, when it has
detected all its predecessors having failed, eventually some replica
thinks it is the head.

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
<{tas(lk)}> & test-and-set on <{!lk}> \\
\hline
<{cas(ptr, old, new)}> & compare-and-swap on <{!ptr}> \\
\hline
<{BinSem(v)}> & return a binary semaphore initialized to <{v}> \\
\hline
<{Lock()}> & return a binary semaphore initialized to <{False}> \\
\hline
<{acquire(bs)}> & acquire binary semaphore <{!bs}> \\
\hline
<{release(bs)}> & release binary semaphore <{!bs}> \\
\hline
<{Condition()}> & return a condition variable \\
\hline
<{wait(c, lk)}> & wait on condition variable <{!c}>
and lock <{lk}> \\ \hline
<{notify(c)}> & notify a thread waiting on condition variable <{!c}> \\
\hline
<{notifyAll(c)}> & notify all threads waiting on condition variable <{!c}> \\
\hline
<{Semaphore(cnt)}> & return a counting semaphore initialized to <{cnt}> \\
\hline
<{P(sema)}> & procure <{!sema}>  \\
\hline
<{V(sema)}> & vacate <{!sema}>  \\
\hline
<{Queue}>() & return a synchronized queue object \\
\hline
<{get(q)}> & return next element of <{q}>, blocking if empty \\
\hline
<{put(q, item)}> & add <{item}> to <{a}> \\
\hline
\end{tabular}
